---
title: "CompareLIs"
author: "Dorothy Bishop"
date: "2023-05-13"
output: html_document
---

18/5/23
updated after adding 95% CI for weighted LI mean
20/5/23
updated to read in the big LI file: li_toolbox_all_tasks_masks_LB.csv
This has 5 and 95% CI
31/5/23
Modified to cope with corrected MCA data, now with binarised mask. NB the raw data files for both toolbox and diff method have been corrected.
7/7/23
To rerun with new outputs from mirror.m and li.m that use 2.5 and 97.5 percentiles
9/3/24
Adjusted for Maria COLA data

NB: FILE LOCATIONS
When you run here(), the default directory for reading and writing becomes the project folder. I have put the raw data files in a folder called maria_2024 within that folder.

If you have other locations, you could set a directory name and then substitute that for cases where you see 'here(maria_2024/' at start of file read

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(here)
require(janitor) #clean up names from external files
require(ggpubr)
require(RVAideMemoire) #for CI for Spearman r
```

## Data from fMRI

6 tasks: Word Generation, Sentence Generation, Phonological Decision, Semantic Comprehension, Word Comprehension, Syntactic Decision (the 'comprehension' tasks also referred to as decision tasks in writeups)

Note from MS: "no participant BAN17 data in the GAM_summary.csv document (their Doppler data was rejected in the GAM analysis), but you will still see their data for the Wilke toolbox and the mirror method (in the files LIs_Wilke_toolbox.csv and flip_allmasks_tasks_feb2024.csv. Their data is omitted in the comparison between fTCD and fMRI when using the script.

Participant OXF13 did not perform the WC task in fMRI. 
We had originally 24 participants in Bangor and 27 in Oxford, without one that was rejected from Bangor data, the final sample size is 50. 


```{r readdata}
demog <- read.csv(here("Maria_2024/GAM_summary.csv"))
# we'll use this just for demographics, handedness etc

# Read in data processed with li.m
allLI <- read.csv(here("Maria_2024/LIs_Wilke_toolbox.csv"))
allLI <- clean_names(allLI) #from janitor package - names more uniform

#need to get an id that will be the same for all files

allLI$id <- substr(allLI$input_image,1,5) #just first 3 chars
idlist <-unique(allLI$id) #complete list of ids - will use later for matching up

#need to code task: this is buried in input_image
allLI$task <- NA
taskID <- c('WG_tstat1','SG_tstat1','PD_tstat1','WC_tstat1','SC_tstat1','SD_tstat1')
task <- c('WG','SG','PD','WC','SC','SD')

for (t in 1:length(task)){
  w<-which(grepl(taskID[t],allLI$input_image)==TRUE)
  allLI$task[w] <-task[t]
}

#now create col for mask
mask <- c('mca','frontal','temporal','parietal')
allLI$mask <- NA
for (m in 1:length(mask)){
  maskname <- mask[m]
  w<-which(grepl(maskname,allLI$inclusive_mask)==TRUE) #find rows with this name
  allLI$mask[w] <-maskname
}

#Now read in mirror files

allmirror <- read.csv(here("Maria_2024/flip_allmasks_tasks_feb2024.csv"))
allmirror<-clean_names(allmirror) #this does some odd things to some names but we will fix later!
allmirror$filename <- allmirror$id
#find the id in the filename
allmirror$id <-NA #initialise
for (i in 1:length(idlist)){
  id <- idlist[i]
  w<-which(grepl(id,allmirror$filename)==TRUE) #find rows with this id in filename
  allmirror$id[w] <-id
}

for (t in 1:length(task)){
  w<-which(grepl(taskID[t],allmirror$filename)==TRUE)
  allmirror$task[w] <-task[t]
}

#Now find the GAM values for doppler
#We have 6 tasks: WG,SG,PD,WC,SC,SD
WG_LI <- read.csv(here("maria_2024/GAM_Summary_WG.csv"))
SG_LI <- read.csv(here("maria_2024/GAM_Summary_SG.csv"))
PD_LI <- read.csv(here("maria_2024/GAM_Summary_PD.csv"))
WC_LI <- read.csv(here("maria_2024/GAM_Summary_WC.csv"))
SC_LI <- read.csv(here("maria_2024/GAM_Summary_SC.csv"))
SD_LI <- read.csv(here("maria_2024/GAM_Summary_SD.csv"))
WG_LI$C_lowCI <- WG_LI$LI.est-1.96*WG_LI$LIest.se
WG_LI$C_hiCI <- WG_LI$LI.est+1.96*WG_LI$LIest.se
SG_LI$C_lowCI <- SG_LI$LI.est-1.96*SG_LI$LIest.se
SG_LI$C_hiCI <- SG_LI$LI.est+1.96*SG_LI$LIest.se
PD_LI$C_lowCI <- PD_LI$LI.est-1.96*PD_LI$LIest.se
PD_LI$C_hiCI <- PD_LI$LI.est+1.96*PD_LI$LIest.se
WC_LI$C_lowCI <- WC_LI$LI.est-1.96*WC_LI$LIest.se
WC_LI$C_hiCI <- WC_LI$LI.est+1.96*WC_LI$LIest.se
SC_LI$C_lowCI <- SC_LI$LI.est-1.96*SC_LI$LIest.se
SC_LI$C_hiCI <- SC_LI$LI.est+1.96*SC_LI$LIest.se
SD_LI$C_lowCI <- SD_LI$LI.est-1.96*SD_LI$LIest.se
SD_LI$C_hiCI <- SD_LI$LI.est+1.96*SD_LI$LIest.se
```

#scale the doppler values by dividing by 6, and the mirror diff values by dividing by 2.7.
#these values determined by looking at max values
#Aim to bring into range with the LIs with scale from -1 to 1

#We will then make a big dataframe with all the data from doppler and fmri bolted together, with columns for id, method, mask, task,LI, lowCI, highCI


```{r makebigdf}
demog$id <- substr(demog$ID,1,5) #again, we can use first 3 digits of ID to uniquely identify participants

#Long form with methods, tasks, masks all stacked

bigdf <- allmirror[,c('id','meanvol_l','meanvol_r','meandiff','diff_low_ci','diff_hi_ci','mask','task')]
names(bigdf)[c(4,5,6)]<-c('mean','lowCI','hiCI') #these are mean and CI for lat index
#scale mirror values by dividing by 2.7 (maximum val for hi CI is 2.689)
bigdf[,4:6]<-bigdf[,4:6]/2.7
bigdf$method <- 'mirror'


#NOW add data from toolbox

#cols for var weighting and clustering just used here as padding for columns with mean L and R activation - will be renamed
addbit<-allLI[,c('id','variance_weighting','clustering','wm_li','wm_low_ci','wm_high_ci','mask','task')]
addbit$method='toolbox'
addbit[,2:3]<-NA
names(addbit)<-names(bigdf)
bigdf<-rbind(bigdf,addbit)


#Now add Doppler - 6 tasks- use NExtreme, nFinal,odd and even for padding
#'Filename','nExtreme','nFinal','C_LI.est','C_lowCI','C_hiCI','odd','even'

dopbit1 <- WG_LI[,c('ID','nExtreme','nFinal','LI.est','C_lowCI','C_hiCI','odd','even')]
dopbit1$even <- 'WG'

dopbit2 <- SG_LI[,c('ID','nExtreme','nFinal','LI.est','C_lowCI','C_hiCI','odd','even')]
dopbit2$even <- 'SG'


dopbit3 <- PD_LI[,c('ID','nExtreme','nFinal','LI.est','C_lowCI','C_hiCI','odd','even')]
dopbit3$even <- 'PD'


dopbit4 <- WC_LI[,c('ID','nExtreme','nFinal','LI.est','C_lowCI','C_hiCI','odd','even')]
dopbit4$even <- 'WC'


dopbit5<- SC_LI[,c('ID','nExtreme','nFinal','LI.est','C_lowCI','C_hiCI','odd','even')]
dopbit5$even <- 'SC'


dopbit6<- SD_LI[,c('ID','nExtreme','nFinal','LI.est','C_lowCI','C_hiCI','odd','even')]
dopbit6$even <- 'SD'

dopbit <- rbind(dopbit1,dopbit2,dopbit3,dopbit4,dopbit5,dopbit6)

dopbit$method <- 'Doppler'
names(dopbit) <- names(bigdf)
dopbit$id <- substr(dopbit$id,1,5)
dopbit[,2:3]<-NA
dopbit$mask <- 'mca'

#Scale doppler by dividing by 6
w<-which(names(dopbit)=='mean')
dopbit[,w:(w+2)]<-dopbit[,w:(w+2)]/6

#now exclude those who did not do fMRI
w<-which(dopbit$id %in% bigdf$id)
bigdf<- rbind(bigdf,dopbit[w,])

bigdf$hand <- NA
bigdf$lat_group <- NA

for (i in 1:nrow(demog)){
  w<-which(bigdf$id==demog$id[i])
  bigdf$hand[w]<-demog$hand_self_report[i]
  bigdf$lat_group[w]<-demog$group_lat[i]
}

#We include a colour code that can be used later in plotting
bigdf$nonRH <- 1
bigdf$handcolour <- 'blue'
bigdf$nonRH[bigdf$hand=="R"] <- 0
bigdf$handcolour[bigdf$hand=="R"]  <-'red' #Red for right-handed


#NB so far the lat_group is Lisa's original grouping
#Here I will add a new lat_group based on whether this measures CI crossed zero

bigdf$nulatgroup <- 'B'
bigdf$nulatgroup[bigdf$lowCI<0 & bigdf$hiCI<0 ] <- 'R'
bigdf$nulatgroup[bigdf$lowCI>0 & bigdf$hiCI>0 ] <- 'L'

#We also want a new DopLatGroup specific to task, where the task has been given
#Just two tasks: PP1 and WG1. This is just used with LvsR plots so just record this for the mirror method, where
#we have the voxel mean vols
bigdf$Doplat<-NA
for (tt in c('WG','SG','PD','WC','SC','SD')){
  myf <- filter(bigdf,bigdf$method=='Doppler',bigdf$task==tt)
  for (n in 1:nrow(myf)){
    w<-intersect(which(bigdf$id==myf$id[n]), which(bigdf$task==tt))
    bigdf$Doplat[w]<-myf$nulatgroup[n]
  }
  
}
write.csv(bigdf,here('maria_2024/bigdf_tstat1.csv'))

```


Now we can look at some plots - start within task

```{r tryplot}
tryplot <- 1 #this flag can be set to zero so we can skip this step if need be
if (tryplot==1){
 
#remove BAN17 as missing data create issues
  bigdf_all<-bigdf #save original
 # bigdf<-bigdf[bigdf$id!='BAN17',] #now remove BAN17 who has missing data that messes up plots

#method options are "Doppler","mirror","toolbox" 
mymethod <- c('Doppler','mirror','toolbox')
#we select these by specifying 1, 2 or 3
methX <- 1
methY <- 2
methodX <- mymethod[methX]
methodY <- mymethod[methY]
methodlabel <- c('Doppler - rescaled','mirror - rescaled','LI toolbox')

allplot <- list() #we'll save the plots in a list so we can flexibly reference them later with ggarrange
mycount <-0
for (thismask in 1){
#for (thismask in 1:3){
mymask <- 'mca'

tlist <- 1:6 # tasks to include: WG,SG,PD,WC,SC,SD
for (t in tlist){
mycount <- mycount+1
mytask <- task[t]
xfile <- filter(bigdf,method==methodX & task==mytask & mask==mymask)
yfile <-  filter(bigdf,method==methodY & task==mytask & mask==mymask)
#Check that rows are equivalent
s<-c(xfile$id,yfile$id)
myn<-table(s)
my2 <- names(which(myn==2))
xfile<-xfile[xfile$id %in% my2,]
yfile <- yfile[yfile$id %in% my2,]
w<-which(names(xfile)=='mean')
names(xfile)[w:(w+2)]<-c('x','xmin','xmax')
names(yfile)[w:(w+2)]<-c('y','ymin','ymax')

forplot<-cbind(xfile,yfile[,w:(w+2)]) #bind x and y columns to make file for plotting

correlcoef <- round(cor.test (forplot$x, forplot$y, method = "spearman",  exact = FALSE)$estimate,3)

corspear <-spearman.ci(forplot$x,forplot$y, nrep = 1000, conf.level = 0.95)
#correlcoef <- paste0('$r_s$ = ',correlcoef,' [',round(corspear$conf.int[1],3),', ',round(corspear$conf.int[2],3),']')
text3 <-paste0('Spearman r: \n',correlcoef,' [',round(corspear$conf.int[1],3),', ',round(corspear$conf.int[2],3),']')

forplot$colgroup= forplot$Doplat #assign for lat_group for colour contrast
if(methX>1 && methY>1){forplot$colgroup <- "1"} #no col contrast if only fMRI (needs to be nonnumeric!)

#out of range CIs mess up plot - set to limits
forplot$xmin[forplot$xmin< (-1)]<- (-1)
forplot$ymin[forplot$ymin< (-1)]<- (-1)
forplot$xmax[forplot$xmax>1]<- 1
forplot$ymax[forplot$ymax>1]<- 1

myplot <- ggplot(data = forplot,aes(x = x,y = y)) + 
  geom_point(aes(col=colgroup),size=.5)+
  geom_errorbar(aes(ymin = ymin,ymax = ymax,col=colgroup)) + 
  geom_errorbarh(aes(xmin = xmin,xmax = xmax,col=colgroup))+
  geom_hline(yintercept = 0, linetype="solid",col='grey') + 
  geom_vline(xintercept = 0, linetype= "solid", col='grey')  + 
  xlab(methodlabel[methX])+
  ylab(methodlabel[methY])+
  lims(x= c(-1,1), y = c(-1,1))+
  ggtitle(paste0(mymask,": ",mytask))+
  annotate("text",x = -.5, y = .9, label = text3,col='black',size=3)+
   theme(legend.position = "none")

mylegend=FALSE
if(methX==1 || methY==1){
myplot <- myplot +
  guides(col=guide_legend(title='Laterality category: fTCD WG'))
  mylegend=TRUE
}

if(methX>1 && methY>1){
myplot <- myplot +
   scale_color_manual(values = "blue") #just use blue points if plotting only fMRI
}

allplot[[mycount]] <- myplot
}
if(length(allplot)==3){
myallplot <- ggarrange(allplot[[1]],allplot[[3]],allplot[[2]],ncol=3,nrow=1,common.legend=mylegend)
savename<-paste0(methodX,'_',methodY,'_3tasks_',mymask,'.jpg')
}
if(length(allplot)==2){
myallplot <- ggarrange(allplot[[1]],allplot[[2]],ncol=2,nrow=1,common.legend=mylegend)
savename<-paste0(methodX,'_',methodY,'_2tasks_',mymask,'.jpg')
}
if(length(allplot)==6){
myallplot <- ggarrange(allplot[[1]],allplot[[2]],allplot[[3]],allplot[[4]],allplot[[5]],allplot[[6]],ncol=2,nrow=3,common.legend=mylegend)
savename<-paste0(methodX,'_',methodY,'_6tasks_',mymask,'.jpg')
}
ggsave(here('maria_2024/Plots',savename),myallplot,width=8,height=9,dpi=300)
}

}
```


Plotting L and R activations from fMRI.
This allows us to see whether L vs R activation switches for cerebellum. (More L activation is points above the line; more R activation is points below the line).
Also allows us to see whether there is deactivation in parietal (mean activation below zero).

```{r LRplot}
LRplot <- 1
if(LRplot==1){
#column with 'diff'
masktitle <- c('mca','frontal','temporal','parietal','cerebellar')
mask<-masktitle
tasktitle <- c('WG','SG','PD','WC','SC','SD')
methodX <- 'mirror'

#We'll save separately for each mask
for (thismask in 1:5){
mymask <- mask[thismask]
mycount <-0
allplot <- list()

for (t in c(1:6)){ #WG, SG, PD, WC, SC, SD
mycount <- mycount+1
mytask <- task[t]
forplot <- filter(bigdf,method==methodX & task==mytask & mask==mymask)
w<-which(is.na(forplot$Doplat)) #find and remove rows with no Doppler lat
if(length(w)>0){
forplot <- forplot[-w,]
}
names(forplot)[2:3]<-c('Left','Right')


#correlcoef <- round(cor.test (forplot$Left, forplot$Right, method = "spearman",  exact = FALSE)$estimate,3) #not used
#censor values at 3
mycensor <- 3
forplot$Left[forplot$Left< (-mycensor)]<- (-mycensor)
forplot$Right[forplot$Right< (-mycensor)]<- (-mycensor)
 forplot$Left[forplot$Left>mycensor]<- mycensor
 forplot$Right[forplot$Right>mycensor]<- mycensor
myplot <- ggplot(data = forplot,aes(x = Right,y = Left)) + 
  geom_point(aes(col=Doplat),size=1.5)+
  geom_hline(yintercept = 0, linetype="solid",col='grey') + 
  geom_vline(xintercept = 0, linetype= "solid", col='grey')  + 
  geom_abline(intercept=0,slope=1,linetype= "dotted", col='black')+
  geom_hline(yintercept = mean(forplot$Left[forplot$Doplat=='L']),col='blue',linetype= "dashed")+
  geom_vline(xintercept = mean(forplot$Right[forplot$Doplat=='L']),col='blue',linetype= "dashed")+
  xlab("Activation R")+
  ylab("Activation L")+
   xlim(-mycensor,mycensor)+
   ylim(-mycensor,mycensor)+
  ggtitle(paste0(masktitle[thismask],":\n",tasktitle[t]))+
  scale_color_manual(values=c( "gray40","blue","red"))+
  guides(col=guide_legend(title='Laterality category: fTCD'))


 
allplot[[mycount]] = myplot
}

myallplot <- ggarrange(allplot[[1]],allplot[[2]],allplot[[3]],allplot[[4]],allplot[[5]],allplot[[6]],ncol=2,nrow=3,common.legend=TRUE)
savename<-paste0('LR_6tasks_',mymask,'.png')
ggsave(here('maria_2024/Plots',savename),myallplot,width=6,height=10,dpi=300)
}
}
```

Here we consider how Doppler activation (coded as MCA region) compares with activation in different ROIs for fMRI

```{r plot5regions}

#method options are "Doppler","mirror","toolbox" - now with mirrorLI which is LI from mirror data
mymethod <- c('Doppler','mirror','toolbox')

tlist<- c('WG','SG','PD','WC','SC','SD')
for (t in 1:length(tlist)){ #new plot started with new task
mytask <- tlist[t]
longtask <- mytask
allplot <- list() #clear list
mycount <-0
methX <- 1 #always compare with Doppler
for (methY in 2){ #methods 2 and 3 will be compared side by side
methodX <- mymethod[methX]
methodY <- mymethod[methY]
methodlabel <- c('fTCD GAM - rescaled','LI from mirror','fMRI: LI toolbox')




for (thismask in 2:4){
mymask <- mask[thismask]
mytitle <- paste0(str_to_title(mymask),": ",longtask)
if(thismask==1){
  mytitle <- paste0(str_to_upper(mymask),": ",longtask)
}

mycount <- mycount+1 #used to count the plots

xfile <- filter(bigdf,method==methodX & task==mytask & mask=='mca') #doppler only has mca equiv
yfile <-  filter(bigdf,method==methodY & task==mytask & mask==mymask)



#Check that rows are equivalent
s<-c(xfile$id,yfile$id)
myn<-table(s)
my2 <- names(which(myn==2))
xfile<-xfile[xfile$id %in% my2,]
yfile <- yfile[yfile$id %in% my2,]
w<-which(names(xfile)=='mean')
names(xfile)[w:(w+2)]<-c('x','xmin','xmax')
names(yfile)[w:(w+2)]<-c('y','ymin','ymax')
pbilat <- round(100*length(which(yfile$nulatgroup=='B'))/nrow(yfile),0) #% bilat
forplot<-cbind(xfile,yfile[,w:(w+2)])

corspear <-spearman.ci(forplot$x,forplot$y, nrep = 1000, conf.level = 0.95)
rs <- round(corspear$estimate,3)
ci1<-round(corspear$conf.int[1],3)
ci2<-round(corspear$conf.int[2],3)
cortext <- paste0('r[s] = ',rs,'\n[',ci1,', ', ci2,']')

#add some text re % bilateral


myplot <- ggplot(data = forplot,aes(x = x,y = y)) + 
  geom_point(aes(col='black'),size=.5)+
  geom_errorbar(aes(ymin = ymin,ymax = ymax,col='black')) + 
  geom_errorbarh(aes(xmin = xmin,xmax = xmax,col='black'))+
  geom_hline(yintercept = 0, linetype="solid",col='grey') + 
  geom_vline(xintercept = 0, linetype= "solid", col='grey')  + 
  xlab(methodlabel[methX])+
  ylab(methodlabel[methY])+
  xlim(-1,1)+
  ylim(-1,1)+
  ggtitle(mytitle)+
  annotate("text",x = -.5, y = .9, label = cortext,col='black')+
  annotate("text",x = 0.65, y = -.85, label = paste0(pbilat,'% bilateral'),col='black')+
  guides(col=guide_legend(title='Laterality category: fTCD WG'))

allplot[[mycount]] = myplot
}
}

myallplot <- ggarrange(allplot[[1]],allplot[[4]],allplot[[2]],allplot[[5]],allplot[[3]],
                       allplot[[6]],nrow=3,ncol=2,common.legend=TRUE)
savename<-paste0('Doppler_fmri2meth_3masks_',mytask,'.jpg')

ggsave(here('maria_2024/Plots',savename),myallplot,width=8,height=15,dpi=300)




}


```

